---
title: "Running Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    source_code: embed
    theme: bootstrap
---

```{r package-loading-and-setup}
library(dplyr)
library(ggplot2)
library(lubridate)
```
 
```{r strava-api}
# Strava API code lifted directly from the link below
# https://bldavies.com/blog/accessing-strava-api/

credentials <- yaml::read_yaml("credentials.yaml")

initiate_oauth_app <- httr::oauth_app(
    appname = "strava",
    key = credentials$client_id,
    secret = credentials$client_secret)

endpoint <- httr::oauth_endpoint(
    request = NULL,
    authorize = "https://www.strava.com/oauth/authorize",
    access = "https://www.strava.com/oauth/token")

token <- httr::oauth2.0_token(
    endpoint = endpoint,
    app = initiate_oauth_app,
    scope = "activity:read_all",
    as_header = FALSE)

df_list <- list()
i <- 1
done <- FALSE

while (!done) {
  req <- httr::GET(
    url = "https://www.strava.com/api/v3/athlete/activities",
    config = token,
    query = list(per_page = 200, page = i)
  )
  df_list[[i]] <- jsonlite::fromJSON(httr::content(req, as = "text"), flatten = TRUE)
  if (length(httr::content(req)) < 200) {
    done <- TRUE
  } else {
    i <- i + 1
  }
}

df <- jsonlite::rbind_pages(df_list)
```

```{r calendar-dates}
lst_years <-
    2022:year(Sys.Date()) |> 
    purrr::map(\(x) 
               tibble::as_tibble_col(
                   x = seq(
                       as_date(glue::glue("{x}-01-01")),
                       as_date(glue::glue("{x}-12-31")), by = "+1 day"),
                   column_name = "date") |> 
               mutate(week_no = isoweek(date)) |> 
               filter(
                   !(month(date) == 1 & week_no == 52),
                   date <= Sys.Date())) |> 
    purrr::set_names(paste0("y", 2022:year(Sys.Date())))

lst_week_breaks <- 
    lst_years |> 
    purrr::map(\(x) 
               x |> 
                   slice_head(n = 1, by = week_no) |> 
                   rename(week_start_date = date))
```

```{r wrangled-running-data}
running_log <- 
    tibble::as_tibble(df) |> 
    filter(
        sport_type == "Run",
        year(start_date) >= 2022) |> 
    mutate(
        run_date = as_date(start_date),
        run_year = if_else(name == "NYRR Midnight Run", year(run_date) - 1, year(run_date)),
        run_time = hms::hms(seconds_to_period(elapsed_time)),
        run_day = wday(run_date, abbr = FALSE, label = TRUE),
        distance_km = round(distance / 1000, 2),
        distance_mile = round(distance_km * 0.621371, 2),
        pace_sec_per_km = round(elapsed_time / distance_km, 2),
        pace_sec_per_mile = round(pace_sec_per_km * 1.609344, 2),
        avg_pace = hms::hms(round(seconds_to_period(pace_sec_per_mile), 0)),
        total_elevation_gain = round(total_elevation_gain * 3.281, 2),
        is_race = if_else(workout_type %in% 1, TRUE, FALSE),
        name = if_else(is_race == TRUE, name, NA)) |> 
    select(
        name,
        is_race,
        run_year,
        run_date,
        run_time,
        run_day,
        distance_mile,
        total_elevation_gain,
        elev_low,
        elev_high,
        average_heartrate,
        run_time,
        avg_pace)
```

```{r official-race-results-table}
official_race_results_tbl <- 
    googlesheets4::read_sheet("1tuC2WBHkDdGDOIbmgnZT32qw7pm4rCXma5Gp7eS41XQ") |> 
    mutate(
        date = as_date(date),
        across(c(official_time, official_pace), function(x) hms::as_hms(x)))

race_schedule <- 
    official_race_results_tbl |> 
    filter(is.na(official_time))
```

```{r official-race-results-metrics}
lst_race_metrics <- 
    2022:year(Sys.Date()) |> 
    purrr::map(\(x) list()) |> 
    purrr::set_names(paste0("y", 2022:year(Sys.Date())))

lst_no_of_races <- 
    purrr::map(
        .x = 2022:year(Sys.Date()),
        .f = function(x) 
            official_race_results_tbl |> 
            filter(year(date) == x & !is.na(official_time)) |> 
            count(name = "no_of_races") |> 
            tibble::deframe()) |> 
    purrr::set_names(paste0("y", 2022:year(Sys.Date())))

lst_no_of_miles <- 
    purrr::map(
        .x = 2022:year(Sys.Date()),
        .f = function(x) 
            official_race_results_tbl |> 
            filter(year(date) == x & !is.na(official_time)) |> 
            summarize(no_of_miles = sum(miles)) |> 
            tibble::deframe()) |> 
    purrr::set_names(paste0("y", 2022:year(Sys.Date())))

lst_avg_paces <- 
    purrr::map(
        .x = 2022:year(Sys.Date()),
        .f = function(x) 
            official_race_results_tbl |> 
            filter(year(date) == x & !is.na(official_time)) |> 
            summarize(avg_pace = mean(official_pace)) |> 
            transmute(avg_pace = hms::hms(round(seconds_to_period(avg_pace), 0))) |> 
            tibble::deframe()) |> 
    purrr::set_names(paste0("y", 2022:year(Sys.Date())))

lst_race_metrics$y2022 <- list(lst_no_of_races$y2022, lst_no_of_miles$y2022, lst_avg_paces$y2022)
lst_race_metrics$y2023 <- list(lst_no_of_races$y2023, lst_no_of_miles$y2023, lst_avg_paces$y2023)
lst_race_metrics$y2024 <- list(lst_no_of_races$y2024, lst_no_of_miles$y2024, lst_avg_paces$y2024)

names(lst_race_metrics$y2022) <- c("no_of_races", "no_of_miles", "avg_pace")
names(lst_race_metrics$y2023) <- c("no_of_races", "no_of_miles", "avg_pace")
names(lst_race_metrics$y2024) <- c("no_of_races", "no_of_miles", "avg_pace")

official_race_results_lst <-
    purrr::map(
        .x = 2022:year(Sys.Date()),
        .f = (\(x) {
            official_race_results_tbl |> 
            filter(year(date) == x) |> 
            mutate(race_status = if_else(is.na(official_time), "Future Race(s)", "Completed Race(s)")) |> 
            select(race_status, date, race, corral, official_time, official_pace, url) |> 
            group_by(race_status) |> 
            arrange(desc(date), .by_group = TRUE) |> 
            DT::datatable(
                rownames = FALSE,
                extensions = 'RowGroup',
                escape = FALSE,
                options = list(
                    rowGroup = list(dataSrc = 0),
                    bPaginate = FALSE,
                    dom = "t",
                    columnDefs = list(
                        list(targets = c(0, 6), visible = FALSE),
                        list(targets = 2, render = DT::JS(
                            "function(data, type, row){
                          if(type === 'display' && row[6] !== null){
                            var a = '<a href=\"' + row[6] + '\">' + data + '</a>';
                            return a;
                          } else {
                            return data;
                          }
                        }")))),
                colnames = c(
                    "Date" = "date",
                    "Race" = "race",
                    "Corral" = "corral",
                    "Time" = "official_time",
                    "Average Pace" = "official_pace"))
        })) |> 
    purrr::set_names(paste0("y", 2022:year(Sys.Date())))
```

```{r weigh-in-data}
weight_hx <- 
    googlesheets4::read_sheet("1Iyn-aMoDoy7MGlqtYJGfBq_7NbjzDArvJMksIMWDpQQ") |> 
    janitor::clean_names() |> 
    select(date_time, weight_lb, bmi)

plt_wgt_tracking_list <-
    purrr::map(
        .x = 2022:year(Sys.Date()),
        .f = function(x) {
            plt <- 
                weight_hx |> 
                filter(between(
                    date_time,
                    as_datetime(glue::glue("{x}-01-01 00:00:00")),
                    as_datetime(glue::glue("{x}-12-31 23:59:59")))) |> 
                ggplot() +
                geom_line(aes(x = date_time, y = weight_lb), colour = "#739dc7") + 
                scale_x_datetime(date_breaks = "1 month", date_labels = "%b %y") + 
                scale_y_continuous("lbs", n.breaks = 12) + 
                theme(
                    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
                    axis.title.x = element_blank())
            
            plotly::ggplotly(plt)}) |> 
    purrr::set_names(paste0("y", 2022:year(Sys.Date())))
```

2024
===================================== 

Row
-------------------------------------

### No. of Miles Ran

```{r total-mileage-2024}
flexdashboard::valueBox(
    running_log |> 
        filter(run_year == 2024) |> 
        summarize(sum(distance_mile)) |> 
        tibble::deframe(),
    icon = "fa-person-running")
```

### Average Run Pace (All Distances)

```{r average-run-pace-2024}
flexdashboard::valueBox(
    running_log |> 
        filter(
            run_year == 2024,
            is_race == FALSE) |> 
        summarize(mean(avg_pace)) |> 
        tibble::deframe() |> 
        hms::as_hms() |> 
        hms::round_hms(digits = 0),
    icon = "fa-gauge-simple")
```

### No. of Races Ran

```{r total-races-ran-2024}
flexdashboard::valueBox(lst_race_metrics[["y2024"]][["no_of_races"]], icon = "fa-medal")
```

### No. of Race Miles Ran

```{r total-race-mileage-2024}
flexdashboard::valueBox(lst_race_metrics[["y2024"]][["no_of_miles"]], icon = "fa-person-running")
```

### Average Race Pace (All Distances)

```{r average-race-pace-2024}
flexdashboard::valueBox(lst_race_metrics[["y2024"]][["avg_pace"]], icon = "fa-gauge-high")
```

Row
-------------------------------------

### **2024 Run Data (excludes races)**
    
```{r run-metrics-table-2024}
running_log |> 
    filter(
        run_year == 2024,
        is_race == FALSE) |> 
    select(run_date, run_time, distance_mile, avg_pace, total_elevation_gain) |> 
    DT::datatable(
        rownames = FALSE,
        escape = FALSE,
        options = list(bPaginate = FALSE, dom = "t"),
        colnames = c(
            "Date" = "run_date",
            "Time" = "run_time",
            "Distance (miles)" = "distance_mile",
            "Elevation Gain (ft)" = "total_elevation_gain",
            "Average Pace" = "avg_pace"))
```

### **2024 Official Race Results**

```{r race-official-data-table-2024}
official_race_results_lst$y2024
```

Row
-------------------------------------

### **2024 Weekly Mileage (includes races)** {.no-padding}

```{r run-mileage-totals-by-week-2024}
plt_weekly_mileage_2024 <- 
    running_log |> 
    filter(run_year == 2024) |> 
    select(run_date, distance_mile) %>%
    left_join(x = lst_years$y2024, y = ., by = c("date" = "run_date")) |> 
    mutate(
        week_no = isoweek(date),
        distance_mile = if_else(is.na(distance_mile), 0, distance_mile)) |> 
    summarize(mileage = sum(distance_mile), .by = week_no) %>%
    left_join(x = ., y = lst_week_breaks$y2024, by = "week_no") |> 
    ggplot() + 
    geom_col(
        aes(
            x = week_start_date,
            y = mileage,
            text = paste0(
                "Week beginning ",
                lst_week_breaks$y2024$week_start_date,
                ": ", mileage, " miles logged")),
        fill = "#aec7e0") + 
    scale_x_date(date_breaks = "1 month", date_labels = "%b %y") + 
    scale_y_continuous(breaks = seq(from = 0, to = 50, by = 2)) + 
    theme(
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        axis.title.x=element_blank()) + 
    ylab("No. of Miles")

plotly::ggplotly(plt_weekly_mileage_2024, tooltip = "text")
```

### **2024 Weight Tracking** {.no-padding}

```{r weight-tracking-2024}
plt_wgt_tracking_list$y2024
```

2023
===================================== 

Row
-------------------------------------

### No. of Miles Ran

```{r total-mileage-2023}
flexdashboard::valueBox(
    running_log |> 
        filter(run_year == 2023) |> 
        summarize(sum(distance_mile)) |> 
        tibble::deframe(),
    icon = "fa-person-running")
```

### Average Run Pace (All Distances)

```{r average-run-pace-2023}
flexdashboard::valueBox(
    running_log |> 
        filter(
            run_year == 2023,
            is_race == FALSE) |> 
        summarize(mean(avg_pace)) |> 
        tibble::deframe() |> 
        hms::as_hms() |> 
        hms::round_hms(digits = 0),
    icon = "fa-gauge-simple")
```

### No. of Races Ran

```{r total-races-ran-2023}
flexdashboard::valueBox(lst_race_metrics[["y2023"]][["no_of_races"]], icon = "fa-medal")
```

### No. of Race Miles Ran

```{r total-race-mileage-2023}
flexdashboard::valueBox(lst_race_metrics[["y2023"]][["no_of_miles"]], icon = "fa-person-running")
```

### Average Race Pace (All Distances)

```{r average-race-pace-2023}
flexdashboard::valueBox(lst_race_metrics[["y2023"]][["avg_pace"]], icon = "fa-gauge-high")
```

Row
-------------------------------------
    
### **2023 Run Data (excludes races)**
    
```{r run-metrics-table-2023}
running_log |> 
    filter(
        run_year == 2023,
        is_race == FALSE) |> 
    select(run_date, run_time, distance_mile, avg_pace, total_elevation_gain) |> 
    DT::datatable(
        rownames = FALSE,
        escape = FALSE,
        options = list(bPaginate = FALSE, dom = "t"),
        colnames = c(
            "Date" = "run_date",
            "Time" = "run_time",
            "Distance (miles)" = "distance_mile",
            "Elevation Gain (ft)" = "total_elevation_gain",
            "Average Pace" = "avg_pace"))
```

### **2023 Official Race Results**
    
```{r race-official-data-table-2023}
official_race_results_lst$y2023
```

Row
-------------------------------------

### **2023 Weekly Mileage (includes races)** {.no-padding}

```{r run-mileage-totals-by-week-2023}
plt_weekly_mileage_2023 <- 
    running_log |> 
    filter(run_year == 2023) |> 
    select(run_date, distance_mile) %>%
    left_join(x = lst_years$y2023, y = ., by = c("date" = "run_date")) |> 
    mutate(
        week_no = isoweek(date),
        distance_mile = if_else(is.na(distance_mile), 0, distance_mile)) |> 
    summarize(mileage = sum(distance_mile), .by = week_no) %>% 
    left_join(x = ., y = lst_week_breaks$y2023, by = "week_no") |> 
    ggplot() + 
    geom_col(
        aes(
            x = week_start_date,
            y = mileage,
            text = paste0(
                "Week beginning ",
                lst_week_breaks$y2023$week_start_date,
                ": ", mileage, " miles logged")),
        fill = "#aec7e0") + 
    scale_x_date(date_breaks = "1 month", date_labels = "%b %y") + 
    scale_y_continuous(breaks = seq(from = 0, to = 50, by = 2)) + 
    theme(
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        axis.title.x=element_blank()) + 
    ylab("No. of Miles")

plotly::ggplotly(plt_weekly_mileage_2023, tooltip = "text")
```

### **2023 Weight Tracking** {.no-padding}

```{r weight-tracking-2023}
plt_wgt_tracking_list$y2023
```

2022
===================================== 

Row
-------------------------------------

### No. of Miles Ran

```{r total-mileage-2022}
flexdashboard::valueBox(
    running_log |> 
        filter(run_year == 2022) |> 
        summarize(sum(distance_mile)) |> 
        tibble::deframe(),
    icon = "fa-person-running")
```

### Average Run Pace (All Distances)

```{r average-run-pace-2022}
flexdashboard::valueBox(
    running_log |> 
        filter(
            run_year == 2022,
            is_race == FALSE) |> 
        summarize(mean(avg_pace)) |> 
        tibble::deframe() |> 
        hms::as_hms() |> 
        hms::round_hms(digits = 0),
    icon = "fa-gauge-simple")
```

### No. of Races Ran

```{r total-races-ran-2022}
flexdashboard::valueBox(lst_race_metrics[["y2022"]][["no_of_races"]], icon = "fa-medal")
```

### No. of Race Miles Ran

```{r total-race-mileage-2022}
flexdashboard::valueBox(lst_race_metrics[["y2022"]][["no_of_miles"]], icon = "fa-person-running")
```

### Average Race Pace (All Distances)

```{r average-race-pace-2022}
flexdashboard::valueBox(lst_race_metrics[["y2022"]][["avg_pace"]], icon = "fa-gauge-high")
```

Row
-------------------------------------
    
### **2022 Run Data (excludes races)**

```{r run-metrics-table-2022}
running_log |> 
    filter(
        run_year == 2022,
        is_race == FALSE) |> 
    select(run_date, run_time, distance_mile, avg_pace, total_elevation_gain) |> 
    DT::datatable(
        rownames = FALSE,
        escape = FALSE,
        options = list(bPaginate = FALSE, dom = "t"),
        colnames = c(
            "Date" = "run_date",
            "Time" = "run_time",
            "Distance (miles)" = "distance_mile",
            "Elevation Gain (ft)" = "total_elevation_gain",
            "Average Pace" = "avg_pace"))
```

### **2022 Official Race Results**
    
```{r race-official-data-table-2022}
official_race_results_lst$y2022
```
   
Row
-------------------------------------

### **2022 Weekly Mileage (includes races)** {.no-padding}

```{r run-mileage-totals-by-week-2022}
plt_weekly_mileage_2022 <- 
    running_log |> 
    filter(run_year == 2022) |> 
    select(run_date, distance_mile) %>% 
    left_join(x = lst_years$y2022, y = ., by = c("date" = "run_date")) |> 
    mutate(
        week_no = isoweek(date),
        distance_mile = if_else(is.na(distance_mile), 0, distance_mile)) |> 
    summarize(mileage = sum(distance_mile), .by = week_no) %>% 
    left_join(x = ., y = lst_week_breaks$y2022, by = "week_no") |> 
    ggplot() + 
    geom_col(
        aes(
            x = week_start_date,
            y = mileage,
            text = paste0(
                "Week beginning ",
                lst_week_breaks$y2022$week_start_date,
                ": ", mileage, " miles logged")),
        fill = "#aec7e0") + 
    scale_x_date(date_breaks = "1 month", date_labels = "%b %y") + 
    scale_y_continuous(breaks = seq(from = 0, to = 50, by = 2)) + 
    theme(
        axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        axis.title.x=element_blank()) + 
    ylab("No. of Miles")

plotly::ggplotly(plt_weekly_mileage_2022, tooltip = "text")
```

### **2022 Weight Tracking** {.no-padding}

```{r weight-tracking-2022}
plt_wgt_tracking_list$y2022
```

NYRR Corrals {data-orientation=columns}
=====================================

Column {data-width=225}
-------------------------------------

```{r nyrr-corral-metrics}
nyrr_best_pace_functional <- as_datetime(paste0("1899-12-30 00:07:50)"))
nyrr_best_pace_pretty <- hms::as_hms(nyrr_best_pace_functional)

nyrr_corral_placement_tbl <- 
    googlesheets4::read_sheet("15iwn0ad1UlaviNjguH2JZq76tycSHvpNlcm10EO19JI") |> 
    select(1:2) |> 
    janitor::clean_names() |> 
    slice(7:17) |> 
    tidyr::unnest(cols = race_distance) |> 
    mutate(
        max_best_pace = race_distance,
        min_best_pace = lag(max_best_pace) + 1,
        best_pace_interval = min_best_pace %--% max_best_pace,
        in_corral = if_else(nyrr_best_pace_functional %within% best_pace_interval, TRUE, FALSE),
        in_corral = if_else(is.na(in_corral), FALSE, in_corral),
        across(min_best_pace:max_best_pace, function(x) hms::as_hms(x))) |> 
    select(corral = pace_units, in_corral, min_best_pace, max_best_pace)
```

### **NYRR Best Pace***

```{r nyrr-best-pace}
flexdashboard::valueBox(nyrr_best_pace_pretty, icon = "fa-gauge-high")
```

### **NYRR Corrals and Current Placement**

```{r nyrr-corral-placement-table}
nyrr_corral_placement_tbl |> 
    DT::datatable(
        rownames = FALSE,
        escape = FALSE,
        options = list(
            bPaginate = FALSE,
            dom = "t",
            columnDefs = list(list(targets = 1, visible = FALSE))),
        colnames = c(
            "Corral" = "corral",
            "Min. Pace" = "min_best_pace",
            "Max. Pace" = "max_best_pace")) |> 
    DT::formatStyle(
        columns = "corral",
        valueColumns = "in_corral",
        target = "row",
        backgroundColor = DT::styleEqual(c(FALSE, TRUE), c("", "#aec7e0")))
```

###

When registering for your first New York Road Runners (NYRR) race you are required to provide a best guess of your finishing time so they can properly seed you. Given that I was only still getting back into running, I entered a projected finish time that ultimately placed me in Corral I.  

On race day--Race to Deliver 4M to Benefit God's Love We Deliver on 2022-11-20--I ended up needing a pit stop during the race and the wait on line for an available Porta Potty added several minutes to my finish time. As such, my first official corral placement, based on available race data, had me in Corral K.  

It's been a steady progression forward since then.

\*Per NYRR's [policy](https://www.nyrr.org/run/guidelines-and-procedures/race-procedures/calculated-pace-and-corral-updates), *"In order to provide a great runner experience at NYRR events, we calculate a best pace using standardized factors that normalize a time run at a distance of 5K or greater to an equivalent 10K time. This allows each runner’s true best performance to be used for corral placement."*  

Column
-------------------------------------

[FUTURE CONTENT COMING HERE SOON. STAY TUNED.]
